---
title: "Assignment 1"
author: "Liam"
date: "9/9/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part 1
## 1. Calculate the following sums.

\begin{align*}
1. S_{1} &= 1 + 2 +...+ 2019
\end{align*}
```{r}
sum(c(1:2019))
```


\begin{align*}
2. S_{2} &= 1^3 + 2^3 +...+ 2019^3
\end{align*}
```{r}
a <- c(1:2019)
sum(a^3)
```


\begin{align*}
3. S_{3} &= 1^1+2^2+3^3+...+2019^{2019}
\end{align*}
```{r}
a <- c(1:2019)
sum(a^a)
```


\begin{align*}
4. S_{4} &= 1^1-2^2+3^3-4^4+...-2018^{2018}+2019^{2019} 
\end{align*}
```{r}
a <- c(1:2019)
b <- a^a
n <- c(1, -1)
sum(b*n)
```


\begin{align*}
5. S_{5} &= 1+1/4+1/9+1/16+1/25+...
\end{align*}
```{r}
a <- c(1:999999)
b <- a^2
sum(a/b)
```


\begin{align*}
6. S_{6} &= 1+\frac{1}{2}+\frac{1}{3}+\frac{1}{4}+...
\end{align*}
```{r}
a <- c(1:999999)
sum(1/a)
```


\begin{align*}
7. S_{7} &= 1+\frac{1}{8}+\frac{1}{27}+\frac{1}{64}+...
\end{align*}
```{r}
a <- c(1:999999)
sum(a/a^3)
```


\begin{align*}
8. S_{8} &= 1-\frac{1}{2}+\frac{1}{3}-\frac{1}{4}+...
\end{align*}
```{r}
a <- c(1:999999)
b <- c(1,-1)
sum(1/(a*b))
```

## 2. The rnorm function generate random variables from normal distribution. Generate a sample of 1000 values from normal distribution with the mean 10 and standard deviation 1.

a. Calculate the mean and standard deviation of the sample.
```{r}
a <- rnorm(1000, mean=10, sd=1)
mean(a)
sd(a)
```

b. Out of 1000 samples, how many do you think are that great than 10? Check your estimation.
```{r}
a <- rnorm(1000, mean=10, sd=1)
a <- a[a>10]
length(a)
```

c. Use hist() function to show the histogram of the sample. 
```{r}
hist(rnorm(1000, mean=10, sd=1))
```

d. Estimate P(X>1) where X~N(2,1)
```{r}
c <- rnorm(10000, mean=2, sd=1)
c <- c[c>1]
length(c)/10000
```

## 3. Consider an experiment of tossing a fair dice.
a. Use the sample (with replacement) function to generate a sample of 1000 values from the experiment.
```{r}
dice <- c(1:6)
a <- sample(dice, 1000, replace=TRUE)
```

b. Calculate the mean and standard deviation of the sample.
```{r}
dice <- c(1:6)
a <- sample(dice, 1000, replace=TRUE)
mean(a)
sd(a)
```


c. How many times the 6 occured?
```{r}
dice <- c(1:6)
a <- sample(dice, 1000, replace=TRUE)
a <- a[a=6]
length(a)
```


d. Use table function to show the frequency of the values.
```{r}
dice <- c(1:6)
a <- sample(dice, 1000, replace=TRUE)
table(a)
```


e. Use prop.table(table()) to show the relative frequency of the values.
```{r}
prop.table(table(a))
```


f. Plot the frequency of the values.
```{r}
plot(prop.table(table(a)))
```


## 4. Consider an experiment of tossing a dice 3 times. Let X1 , X2, and X3 be the number of tossing each dice
a. P(X1 > X2 + X3)
```{r}
x1 <- sample(dice, 1000, replace = TRUE)
x2 <- sample(dice, 1000, replace = TRUE)
x3 <- sample(dice, 1000, replace = TRUE)
sum(x1>x2+x3)/1000
```


b. P(X1^2 > X2 + X3)
```{r}
x1 <- sample(dice, 1000, replace = TRUE)
x2 <- sample(dice, 1000, replace = TRUE)
x3 <- sample(dice, 1000, replace = TRUE)
sum(x1^2>x2+x3)/1000
```


## 5. Using simulation, estimate the probability of getting three tails in a row when tossing a coin 3 times. Hint: one way is to generate a matrix with three columns where each rows is an observation of tossing a coin three times.
```{r}
coin <-c(1:2)
f1 <-sample(coin, 1000, replace=TRUE)
f2 <-sample(coin, 1000, replace=TRUE)
f3 <-sample(coin, 1000, replace=TRUE)
m <-data.frame(f1,f2,f3)
j <-rowSums(m)
sum(j==6)/1000
```


## 6. (Extra Credits/Optional) Using simulation, estimate the probability of getting three tails in a row when tossing a coin 10 times.
```{r}

```


## 7. Central Limit Theorem (CLT). The CLT said that the mean of a sample of a distribution A (no matter what A is) follows normal distribution with the same mean as A. Following the below steps to confim the CLT when A is uniform distribution.
a. Generate 100 samples of uniform distibution from 0 to 1. Each sample has 1000 observations. Use the runif function to do this.
```{r}
s<-100
obs<-1000
a <-matrix(runif(s*obs,0,1),ncol=s)
```

b. Compute the means of the 100 samples. Create vector x containing these means. Hint: You want to put all the samples in a matrix and use rowSums or colSums function.
```{r}
b <-rowSums(a)
x <- b/1000
```

c. By CLT, x must follow normal distribution. Check this by plotting the histogram of x. Does it look like normal distribution? Use hist(x) to plot the histogram of x.
```{r}
hist(x)
```

d.Increase the number (100 and 1000) to see if the distribution of x looks more like normal distribution.
```{r}
s<-1000
obs<-1000
a <-matrix(runif(s*obs,0,1),ncol=s)
b <-rowSums(a)
x <- b/1000
hist(x)
```

e.Try the same procedure with two other distributions for A.
```{r}
#poisson
s<-1000
obs<-1000
a <-matrix(rpois(s*obs,0.5),ncol=s)
b <-rowSums(a)
x <- b/1000
hist(x)

#geometric
s<-1000
obs<-1000
a <-matrix(rgeom(s*obs,0.5),ncol=s)
b <-rowSums(a)
x <- b/1000
hist(x)
```


# Part 2
## 7. Use read.csv function to read in the titanic dataset. You can find the dataset on Blackboard or at Kaggle.com. Use str function to see a summary of the data.
```{r}
titanic<-read.csv("C:\\Users\\student\\Desktop\\Fall2019\\R\\titanic.csv")
str(titanic)
```

## 8. Use knitr::kable function to nicely print out the first 10 rows of the data in markdown.
```{r}
knitr::kable(head(titanic))
```

## 9. Use is.na function and sum function to count the total number of missing values in the data. Count the number of missing values in each columns.
```{r}
colSums(is.na(titanic))
```

## 10. Calculate the average Age of the passengers. You may want to use the parameter na.rm = TRUE in the function mean
```{r}
ageavg<-colMeans(titanic['Age'], na.rm=TRUE)
```

## 11. Replace the missing values of age by the average age calculated previously.
```{r}
titanic[is.na(titanic['Age']), 'Age'] <- ageavg
knitr::kable(head(titanic))
```

## 12. Remove columns Name, PassengerID, Ticket, and Cabin.
```{r}
titanic <- subset(titanic, select = -c(Name,PassengerId,Ticket,Cabin))
knitr::kable(head(titanic))
```

## 13. Calculate the mean age of female passengers
```{r}
mean(titanic$Age[titanic$Sex=='female'])
```

## 14. Calculate the median fare of the passengers in Class 1
```{r}
median(titanic$Fare[titanic$Pclass==1])
```

## 15. Calculate the median fare of the female passengers that are not in Class 1
```{r}
median(titanic$Fare[titanic$Pclass!=1 & titanic$Sex=='female'])
```

## 16. Calculate the median age of survived passengers who are female and Class 1 or Class 2,
```{r}
median(titanic$Age[(titanic$Pclass==1| titanic$Pclass==2) & titanic$Sex=='female' & titanic$Survived==1])
```

## 17. Calculate the mean fare of female teenagers survived passengers
```{r}
mean(titanic$Fare[titanic$Sex=='female' & titanic$Age>12 & titanic$Age<20 & titanic$Survived == 1])
```

## 18. Calculate the mean fare of female teenagers survived passengers for each class
```{r}
titanic2<-subset(titanic, Sex=='female' & titanic$Age>12 & titanic$Age<20 & titanic$Survived == 1)
aggregate(titanic2[,'Fare'], list(titanic2$Pclass),mean)

```

## 19. Calculate the ratio of Survived and not Survived for passengers who are who pays more than the average fare
```{r}
avgFare<-mean(titanic$Fare)
titanic3<-subset(titanic, Fare > avgFare)
print("Survived")
sum(titanic3$Survived==1)/sum(titanic3$Survived==1 | titanic3$Survived==0)
print("Did Not Survive")
sum(titanic3$Survived==0)/sum(titanic3$Survived==1 | titanic3$Survived==0)
```

## 20. Add column that standardizes the fare (subtract the mean and divide by standard deviation) and name it sfare
```{r}
sdFare<-sd(titanic$Fare)
titanic$sfare<-(titanic$Fare-avgFare)/sdFare
head(titanic)
```

## 21. Add categorical variable named cfare that takes value cheap for passengers paying less the average fare and takes value expensive for passengers paying more than the average fare.
```{r}
titanic$cfare <- ifelse(titanic$Fare < avgFare, "cheap", 'expensive')
```

## 22. Add categorical variable named cage that takes value 0 for age 0-10, 1 for age 10-20, 2 for age 20-30, and so on
```{r}
titanic$cage <-memisc::cases('0'=titanic$Age<10,
                    '1'=titanic$Age<20,
                    '2'=titanic$Age<30,
                    '3'=titanic$Age<30,
                    '4'=titanic$Age<40,
                    '5'=titanic$Age<50,
                    '6'=titanic$Age<60,
                    '7'=titanic$Age<70,
                    '8'=titanic$Age<80,
                    '9'=titanic$Age<90,
                    '10'=TRUE)
```

## 23. Show the frequency of Ports of Embarkation. It appears that there are two missing values in the Embarked variable. Assign the most frequent port to the missing ports. Hint: Use the levels function to modify the categories of categorical variables.
```{r}
summary(titanic$Embarked)
titanic$Embarked[titanic$Embarked==""] <- "S"
summary(titanic$Embarked)
```

```{r, eval= FALSE}
rmarkdown::github_document
```